---
title: "simulation_notebook"
author: "yuqimiao"
date: "2022-09-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
library(tidyverse)
# need to connect to cluster
```

## 220930 4 clusters

**Basic setting: **

The setting is to have 4 clusters in total. Dataset 1 separates clusters (1+2) from 3, and 4. Dataset 2 separates clusters 1, 2 and (3+4). That is, using dataset 1 or 2 only, you can detect 3 clusters, although 2 sets of three different clusters. Only when you use both datasets 1 and 2, you can detect 4 clusters. This way, the overlapping information of the 2 datasets are, clusters (1+2) from clusters (3+4). The unique information of dataset 1 is to separate cluster 1 and 2. The unique information of dataset 2 is to separate cluster 3 and 4.

**Parameters: **

* n_feat1 = 1000
* n_feat2 = 100000
* mu1 = c(0, 0, 1, -1)
* mu2 = c(1,-1,0,0)

```{r}
# data read in
dir = "/Volumes/sw2206/yuqi/simu_220930"
files = list.files(dir)
tib_all = NULL
for(i in 1:length(files)){
  tib = readRDS(paste(dir,"/", files[i], sep = ""))
  tib_all = rbind(tib_all, tib)
}

tib_all %>% unnest(res_tib) %>% 
  group_by(noise_sd, kernel, method) %>% 
  summarize(mean_nmi = mean(nmi),
         sd_nmi = sd(nmi)) %>% 
  ggplot(aes(x = noise_sd, y = mean_nmi, color = method))+
  geom_line(aes(linetype=kernel, color = method))+
  # geom_errorbar(aes(ymin=mean_nmi-sd_nmi, ymax=mean_nmi+sd_nmi), width=.2)
  geom_point()
```

2 Potential problems:

1. Parition-cimlr using the true number of clusters for both single data and integrated data, while SNF and cimlr only takes true number of clusters for the integrated data

