---
title: "TCGA_data_basics"
author: "yuqimiao"
date: "2022-09-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

data dimension collection

see: 
Dimension check: 

```{r}
"/Users/miaoyuqi/研究/Shuang project/multiomics_integation/code/weekly_analysis_script/Week2022-03-10/TCGA_data_check.R"
```

Common sample from fitting-distance_list:
```{r}
"./code/weekly_analysis_script/Week2022-09-08/Survival_TCGA.R"
```



# BRCA

* Gene expression: 
  * cancer: 17002  1093
  * normal: 17002   109
* Methylation:
  * cancer: 311121    604
  * normal: 311121     90
* Mutation: 14164   604

* Common sample: 603


# KIRP

* Gene expression: 
  * cancer: 16729   290
  * normal: 16729    30
* Methylation:
  * cancer: 302734    137
  * normal: 302734     45
* Mutation: 8475  137

* Common sample: 137

# lihc

* Gene expression: 
  * cancer: 16085   367
  * normal: 16085    49
* Methylation:
  * cancer: 302747    161
  * normal: 302747     50
* Mutation: 11448   162

* Common sample: 161
